---
title: "Primary analysis"
output: html_document
date: '2022-10-18'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
library(stringr)
library(geepack)
library(ggpubr)
library(cowplot)
library(car)
`%notin%` = Negate(`%in%`)

setwd("~/Valentine/Final Data")

########### 
# Table 1
###########

library(table1)
```



```{r include=FALSE}
plot_multi_histogram <- function(df, feature, label_column) {
  plt <- ggplot(df, aes(x=eval(parse(text=feature)), fill=eval(parse(text=label_column)))) +
    geom_histogram(alpha=0.7, position="identity", aes(y = ..density..), color="black") +
    geom_density(alpha=0.7) +
    # geom_vline(aes(xintercept=mean(eval(parse(text=feature)))), color="black", linetype="dashed", size=1) +
    labs(x=feature, y = "Density")
  plt + guides(fill=guide_legend(title=label_column))
}

StudyParticipants <- read_csv("StudyParticipants 20221025.csv", 
    col_types = cols(TimeEnrollment_to_Exit_days = col_double(),
                     IsExit = col_factor(levels = c("0",  "1")),
                     IsWithdrawn = col_factor(levels = c("0",  "1")),))
# 
# StudyParticipants <- read_csv("StudyParticipants.csv", 
#                               col_types = cols(# CardiacRehabStartDate_ET = col_date(format = "%Y-%m-%d"), 
#                                 # ConsentDate_ET = col_date(format = "%Y-%m-%d"), 
#                                 # EnrollmentDate_ET = col_date(format = "%Y-%m-%d"), 
#                                 # EnrollmentDate_ET_Add_6mo = col_date(format = "%Y-%m-%d"), 
#                                 # WithdrawalDate_ET = col_date(format = "%Y-%m-%d"), 
#                                 # TerminationDate_ET = col_date(format = "%Y-%m-%d"), 
#                                 ExitDate_ET = col_date(format = "%m/%d/%y"), 
#                                 IsExit = col_factor(levels = c("0",  "1")),
#                                 # Time_CardiacRehabStart_to_Withdrawal_days = col_double(), 
#                                 # Time_CardiacRehabStart_to_Termination_days = col_double(), 
#                                 # Time_CardiacRehabStart_to_Exit_days = col_double(), 
#                                 AgeWithdrawal_years = col_double(),
#                                 AgeEnrollment_years = col_double(),
#                                 AgeTermination_years = col_double(), 
#                                 AgeExit_years = col_double()))

StudyParticipants$Age_cat = ifelse(StudyParticipants$AgeEnrollment_years<65, "young", "senior")
StudyParticipants$Age_cat_id = ifelse(StudyParticipants$AgeEnrollment_years<65, -1, 1)
```

### Table 1


```{r include=FALSE}
pvalue <- function(x, ...) {
    # Construct vectors of data y, and groups (strata) g
    y <- unlist(x)
    g <- factor(rep(1:length(x), times=sapply(x, length)))
    if (is.numeric(y)) {
        # For numeric variables, perform a standard 2-sample t-test
        p <- t.test(y ~ g)$p.value
    } else {
        # For categorical variables, perform a chi-squared test of independence
        p <- chisq.test(table(y, g))$p.value
    }
    # Format the p-value, using an HTML entity for the less-than sign.
    # The initial empty string places the output on the line below the variable label.
    c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
}
```



```{r echo=FALSE, warning=FALSE}
StudyParticipants$BMI = as.numeric(StudyParticipants$BMI)


table1(~Gender + Age_cat+Race+Ethnicity+ Indication +Device+CardiacRehabCenter + IsExit+IsWithdrawn+ ComorbidCAD_bool+ComorbidHF_bool+
         ComorbidPCICABG_bool+ComorbidValve_bool+ TimeCardiacRehabStart_to_Enrollment_days+AgeEnrollment_years+TimeEnrollment_to_Exit_days+BMI+BMI_Classification|Cohort, data = StudyParticipants , overall=F, extra.col=list(`P-value`=pvalue) )
```



### Withdrew analysis

```{r echo=FALSE}
## withdrew

table1(~Cohort  +Device + TimeEnrollment_to_Exit_days, data = StudyParticipants[StudyParticipants$IsWithdrawn==1,] )

StudyParticipants[StudyParticipants$IsWithdrawn==1,]$ParticipantIdentifier

```



### Average daily step count

```{r include=FALSE}
library(readr)
Valentine_Primary_Analysis_Data <-  read_csv("Functional Endpoints 20221017.csv", 
    col_types = cols(Date_ET = col_date(format = "%Y-%m-%d"), 
        WearTimeAgg_hours = col_double(), 
        WearTime_hours = col_number(), DailyWearTime_hours = col_number(), 
        ExerciseTimeAgg_min = col_double(), 
        DailyExerciseTime_min = col_double(), 
        StepsAgg = col_double(), DailySteps = col_double(), 
        WalkDistance_m = col_double())) 


Valentine_Primary_Analysis_Data <- Valentine_Primary_Analysis_Data %>%
  arrange(ParticipantIdentifier, Date_ET)%>%
  distinct(ParticipantIdentifier, Date_ET,.keep_all = TRUE)

Valentine_Primary_Analysis_Data$WearTime_hours = if_else(is.na(Valentine_Primary_Analysis_Data$WearTimeAgg_hours),Valentine_Primary_Analysis_Data$WearTime_hours,Valentine_Primary_Analysis_Data$WearTimeAgg_hours)


Valentine_Primary_Analysis_Data$StepsAgg = if_else(is.na(Valentine_Primary_Analysis_Data$StepsAgg),Valentine_Primary_Analysis_Data$DailySteps,Valentine_Primary_Analysis_Data$StepsAgg)

load("Data_step_3month.RData")

Primary_Analysis_Data_summary = Valentine_Primary_Analysis_Data %>% 
  group_by(ParticipantIdentifier, Phase)%>%
  summarise(days_record = sum(StepsAgg>100,na.rm = TRUE),weekly_average = mean(StepsAgg,na.rm = TRUE)) %>%
  na_if("NaN")%>% left_join(StudyParticipants[,c("ParticipantIdentifier","Cohort","Device")])%>%
  rbind(Data_step_3month)%>%
  left_join(StudyParticipants[,c("ParticipantIdentifier","IsWithdrawn")])


# write.csv(Primary_Analysis_Data_summary, file = "step_count.csv")

# length(unique(Primary_Analysis_Data_summary$ParticipantIdentifier))

# Less_3_6 = Primary_Analysis_Data_summary %>% filter(Phase !="3-mo.") %>% filter(days_record < 3 |IsWithdrawn==1 )
# ID_6 = unique(Less_3_6$ParticipantIdentifier)
# 
# Less_3_3 = Primary_Analysis_Data_summary %>% filter(Phase !="6-mo.") %>% filter(days_record < 3 |IsWithdrawn==1 )
# ID_3 = unique(Less_3_3$ParticipantIdentifier)


Less_3_0 = Primary_Analysis_Data_summary %>% filter(Phase =="0-mo.") %>% filter(days_record < 3 )
ID_0 = unique(Less_3_0$ParticipantIdentifier)

Less_3_6 = Primary_Analysis_Data_summary %>% filter(Phase=="6-mo.") %>% filter(days_record < 3 )
ID_6 = unique(Less_3_6$ParticipantIdentifier)

Less_3_3 = Primary_Analysis_Data_summary %>% filter(Phase =="3-mo.") %>% filter(days_record < 3 )
ID_3 = unique(Less_3_3$ParticipantIdentifier)


Total_summary = Primary_Analysis_Data_summary %>%
  select(-days_record)%>%
  pivot_wider(names_from = Phase, values_from = weekly_average)

Total_summary[Total_summary$ParticipantIdentifier %in% ID_0,]$`0-mo.` = NA
Total_summary[Total_summary$ParticipantIdentifier %in% ID_6,]$`6-mo.` = NA
Total_summary[Total_summary$ParticipantIdentifier %in% ID_3,]$`3-mo.` = NA

# %>%filter(ParticipantIdentifier %notin% ID)

Total_summary$weekly_average_dif = Total_summary$`6-mo.` - Total_summary$`0-mo.`

Total_summary$weekly_average_dif3 = Total_summary$`3-mo.`- Total_summary$`0-mo.`


Total_summary = left_join(Total_summary, StudyParticipants[,c("ParticipantIdentifier","CardiacRehabCenter","IsExit","Gender", "Age_cat","Age_cat_id", "Race", "Indication",  "ComorbidCAD_bool",                                 
                                                    "ComorbidHF_bool", "ComorbidPCICABG_bool","ComorbidValve_bool")], by = "ParticipantIdentifier")

Total_summary$Cohort = factor(Total_summary$Cohort, levels = c("Usual","Telehealth"))
Total_summary$Cohort = relevel(Total_summary$Cohort, ref = "Usual")
```


### Comparing the first week and the 3-mo. & 6-mo. average daily step counts

```{r echo=FALSE, warning=FALSE}
#### In the paragraph

plot_multi_histogram(Primary_Analysis_Data_summary, "weekly_average", "Phase")

# table1(~weekly_average_dif+weekly_average_dif3+ `0-mo.` + `3-mo.` + `6-mo.` | Cohort, data = Total_summary, overall=F, extra.col=list(`P-value`=pvalue))

# Stepcount_summary = Primary_Analysis_Data_summary %>%
#   group_by(Cohort, Phase)%>%
#   summarise(step_count_mean = mean(weekly_average,na.rm = T),
#             step_count_sd = sd(weekly_average,na.rm = T),
#             n = sum(!is.na(weekly_average)))




# ## Case
# apply(Total_summary[Total_summary$Cohort == "Telehealth",c(5,6,7)], 2, mean, na.rm = T)
# apply(Total_summary[Total_summary$Cohort == "Telehealth",c(5,6,7)], 2, sd, na.rm = T)
# apply(!is.na(Total_summary[Total_summary$Cohort == "Telehealth",c(5,6,7)]), 2, sum)
# ## Control
# apply(Total_summary[Total_summary$Cohort == "Usual",c(5,6,7)], 2, mean, na.rm = T)
# apply(Total_summary[Total_summary$Cohort == "Usual",c(5,6,7)], 2, sd, na.rm = T)
# apply(!is.na(Total_summary[Total_summary$Cohort == "Usual",c(5,6,7)]), 2, sum)
# 

################ Supplement


# ## Case Apple
# apply(Total_summary[Total_summary$Cohort == "Telehealth"& Total_summary$Device == "Apple Watch",c(5,6,7)], 2, mean, na.rm = T)
# apply(Total_summary[Total_summary$Cohort == "Telehealth"& Total_summary$Device == "Apple Watch",c(5,6,7)], 2, sd, na.rm = T)
# apply(!is.na(Total_summary[Total_summary$Cohort == "Telehealth"& Total_summary$Device == "Apple Watch",c(5,6,7)]), 2, sum)
# 
# 
# ## Case Fitbit
# apply(Total_summary[Total_summary$Cohort == "Telehealth"& Total_summary$Device == "Fitbit",c(5,6,7)], 2, mean, na.rm = T)
# apply(Total_summary[Total_summary$Cohort == "Telehealth"& Total_summary$Device == "Fitbit",c(5,6,7)], 2, sd, na.rm = T)
# apply(!is.na(Total_summary[Total_summary$Cohort == "Telehealth"& Total_summary$Device == "Fitbit",c(5,6,7)]), 2, sum)
# 

# 
# ## Control Apple
# #
# apply(Total_summary[Total_summary$Cohort == "Usual"& Total_summary$Device == "Apple Watch",c(5,6,7)], 2, mean, na.rm = T)
# apply(Total_summary[Total_summary$Cohort == "Usual"& Total_summary$Device == "Apple Watch",c(5,6,7)], 2, sd, na.rm = T)
# apply(!is.na(Total_summary[Total_summary$Cohort == "Usual"& Total_summary$Device == "Apple Watch",c(5,6,7)]), 2, sum)

#
#
# ## Control Fibit
#
# apply(Total_summary[Total_summary$Cohort == "Usual"& Total_summary$Device == "Fitbit",c(5,6,7)], 2, mean, na.rm = T)
# apply(Total_summary[Total_summary$Cohort == "Usual"& Total_summary$Device == "Fitbit",c(5,6,7)], 2, sd, na.rm = T)
# apply(!is.na(Total_summary[Total_summary$Cohort == "Usual"& Total_summary$Device == "Fitbit",c(5,6,7)]), 2, sum)

# Primary_Analysis_Data_summary$Cohort = if_else(Primary_Analysis_Data_summary$Cohort == "Usual", "Control", "Intervention")

p_step_1 = ggplot(Primary_Analysis_Data_summary[Primary_Analysis_Data_summary$Phase!="3-mo.",], 
                aes(x = Phase, y = weekly_average, fill = Cohort))+
  geom_boxplot()+
  theme_bw()+
  facet_wrap(vars(Device))+
  ylab("Weekly Average Step Counts (steps)")

p_step_1

p_step_2 = ggplot(Primary_Analysis_Data_summary[Primary_Analysis_Data_summary$Phase!="6-mo.",], 
                aes(x = Phase, y = weekly_average, fill = Cohort))+
  geom_boxplot()+
  theme_bw()+
  facet_wrap(vars(Device))+
  ylab("Weekly Average Step Counts (steps)")

p_step_2


p_step_3 = ggplot(Primary_Analysis_Data_summary, 
                aes(x = Phase, y = weekly_average, fill = Cohort))+
  geom_boxplot()+
  theme_bw()+
  facet_wrap(vars(Device))+
  ylab("Weekly Average Step Counts (steps)")

p_step_3

```

### Linear model testing the primiary hypothesis

```{r echo=FALSE}

# cat("---------------------------------","\n")
# 
# cat("There are",sum(is.na(Total_summary$`6-mo.`)), "people without last week's average step count, and their IDs are",Total_summary$ParticipantIdentifier[is.na(Total_summary$`6-mo.`)],"\n")
# 
# cat("---------------------------------","\n")
# 
# cat("There are", sum(is.na(Total_summary$`0-mo.`)), "people without first week's step counts, and their IDs are",Total_summary$ParticipantIdentifier[is.na(Total_summary$`0-mo.`)],"\n")
    


cat("---------------------------------","\n")

fit = lm(weekly_average_dif  ~ Cohort * Device, data =Total_summary )
summary(fit)

cat("---------------------------------","\n")

cat("In this model, there are in total",dim(fit[["model"]])[1], "observations used", "\n")

cat("---------------------------------","\n")



linearHypothesis(fit,c("CohortTelehealth = 0", "CohortTelehealth + CohortTelehealth:DeviceFitbit = 0")) 
```

```{r echo=FALSE}


# cat("---------------------------------","\n")
# 
# cat("There are",sum(is.na(Total_summary$weekly_average_last)), "people without last week's average step count, and their IDs are",Total_summary$ParticipantIdentifier[is.na(Total_summary$weekly_average_last)],"\n")
# 
# cat("---------------------------------","\n")
# 
# cat("There are", sum(is.na(Total_summary$weekly_average_first)), "people without first week's step counts, and their IDs are",Total_summary$ParticipantIdentifier[is.na(Total_summary$weekly_average_first)],"\n")
#     
# 
# 
# cat("---------------------------------","\n")

fit = lm(weekly_average_dif3  ~ Cohort * Device, data =Total_summary )
summary(fit)

cat("---------------------------------","\n")

cat("In this model, there are in total",dim(fit[["model"]])[1], "observations used", "\n")

cat("---------------------------------","\n")



linearHypothesis(fit,c("CohortTelehealth = 0", "CohortTelehealth + CohortTelehealth:DeviceFitbit = 0")) 
```
### Sub group analysis

#### Age cat

For Apple Watch

```{r echo=FALSE}
fit_6 = lm(weekly_average_dif ~ Cohort *Age_cat, data = Total_summary[Total_summary$Device=="Apple Watch",] )
summary(fit_6)

```


For Fitbit

```{r echo=FALSE}
fit_6 = lm(weekly_average_dif ~ Cohort *Age_cat, data = Total_summary[Total_summary$Device=="Fitbit",] )
summary(fit_6)

```

#### Gender

For Apple Watch

```{r echo=FALSE}
fit_6 = lm(weekly_average_dif ~ Cohort *Gender, data = Total_summary[Total_summary$Device=="Apple Watch",] )
summary(fit_6)
```


For Fitbit

```{r echo=FALSE}
fit_6 = lm(weekly_average_dif ~ Cohort *Gender, data = Total_summary[Total_summary$Device=="Fitbit",] )
summary(fit_6)
```

#### Presence of Heart Failure

For Apple Watch

```{r echo=FALSE}
fit_6 = lm(weekly_average_dif ~ Cohort *ComorbidHF_bool , data = Total_summary[Total_summary$Device=="Apple Watch",] )
summary(fit_6)
```


For Fitbit

```{r echo=FALSE}
fit_6 = lm(weekly_average_dif ~ Cohort *ComorbidHF_bool, data = Total_summary[Total_summary$Device=="Fitbit",] )
summary(fit_6)
```

#### Cardiac Rehab Center

For Apple Watch

```{r echo=FALSE}
fit_6 = lm(weekly_average_dif ~ Cohort *CardiacRehabCenter, data = Total_summary[Total_summary$Device=="Apple Watch",] )
summary(fit_6)
```


For Fitbit

```{r echo=FALSE}
fit_6 = lm(weekly_average_dif ~ Cohort *CardiacRehabCenter, data = Total_summary[Total_summary$Device=="Fitbit",] )
summary(fit_6)
```

### Survey Results

```{r warning=FALSE, include=FALSE}

SurveyResults <- read_csv("SurveyResults EQ5D 20221031.csv")

baseline_survey = SurveyResults %>% filter(Phase == "0-mo.")%>% select(-c(6:14))
colnames(baseline_survey)[7] =  "Value_baseline"
sixmonth_survey = SurveyResults %>% filter(Phase == "6-mo.")%>% select(-c(6:14))
colnames(sixmonth_survey)[7] =  "Value_6month"




IDs = unique(StudyParticipants$ParticipantIdentifier)
df_baseline = baseline_survey %>% filter(ScoreCategory=="EQ-5D_Health_Scale" ) %>% group_by(ParticipantIdentifier)%>% count() # 219
df_sixmonth = sixmonth_survey %>% filter(ScoreCategory=="EQ-5D_Health_Scale" ) %>% group_by(ParticipantIdentifier)%>% count() # 190


Survey = full_join(baseline_survey,sixmonth_survey, by = intersect(colnames(baseline_survey),colnames(sixmonth_survey)))%>%
  # arrange(ParticipantIdentifier,SurveyName,ScoreType)%>%
  mutate(Value_dif =Value_6month-Value_baseline ) %>%
  filter(ScoreCategory=="EQ-5D_Health_Scale" ) %>%
  distinct(ParticipantIdentifier,.keep_all = T) %>%
  left_join(.,StudyParticipants[,c("ParticipantIdentifier","Cohort","CardiacRehabCenter","IsExit","Gender","Device", "Age_cat","Age_cat_id", "Race", "Indication",  "ComorbidCAD_bool",                                 
                                 "ComorbidHF_bool", "ComorbidPCICABG_bool","ComorbidValve_bool")], by = "ParticipantIdentifier")


Survey$Cohort = factor(Survey$Cohort, levels = c("Usual","Telehealth"))
Survey$Cohort = relevel(Survey$Cohort, ref = "Usual")


```


```{r echo=FALSE}
Survey %>% group_by(Cohort) %>% summarise(baseline_mean = mean(Value_baseline,na.rm = TRUE),
                                          baseline_sd = sd(Value_baseline,na.rm = TRUE),
                                          sixmonth_mean = mean(Value_6month,na.rm = TRUE),
                                          sixmonth_sd = sd(Value_6month,na.rm = TRUE))


cat("ID", setdiff(IDs, df_baseline$ParticipantIdentifier), "don't have baseline EQ-5D Health Scale score, and ID", setdiff(IDs, df_sixmonth$ParticipantIdentifier), "don't have 6month EQ-5D Health Scale score.")

t_test = with(Survey,
     t.test(Value_dif[Cohort == "Telehealth"],Value_dif[Cohort == "Usual"])) #0.4315

print(t_test)

```



### 6MW Distance

```{r include=FALSE}

Valentine_Primary_Analysis_6MW_Data <-  read_csv("SurveyResults 6MW 20221020.csv", 
    col_types = cols(Distance_m = col_double(), 
                     TaskInsertedDate_ET = col_date(format = "%m/%d/%y"), 
        StepStartDate_ET = col_date(format = "%m/%d/%y"), 
        Duration_min = col_double(), SampleDistance_m = col_double(), 
        SampleDuration_min = col_double()))%>%
  left_join(StudyParticipants[,c("ParticipantIdentifier","Cohort")])

# Valentine_Primary_Analysis_6MW_Data = Valentine_Primary_Analysis_6MW_Data  
#


month_0 = Valentine_Primary_Analysis_6MW_Data %>% filter(TaskIdentifier == "6 Minute Walk (0-mo.)")
ref_month0 = month_0 %>%
  group_by(ParticipantIdentifier)%>%
  summarise(TaskInsertedDate = min(TaskInsertedDate_ET),
            StepStartDate_ET = max(StepStartDate_ET,na.rm = T))
ref_month0[sapply(ref_month0, is.infinite)] <- NA
month_0 = left_join(month_0,ref_month0, by = c("ParticipantIdentifier","StepStartDate_ET"))%>%
   mutate(days_complete_task = StepStartDate_ET - TaskInsertedDate)%>%
   filter(days_complete_task<=30) %>%   
   filter(IsDistanceAtLeast100m==1) 
  
month_3 = Valentine_Primary_Analysis_6MW_Data %>% filter(TaskIdentifier == "6 Minute Walk (3-mo.)")
ref_month3 = month_3 %>%
  group_by(ParticipantIdentifier)%>%
  summarise(TaskInsertedDate = min(TaskInsertedDate_ET),
            StepStartDate_ET = max(StepStartDate_ET,na.rm = T))
ref_month3[sapply(ref_month3, is.infinite)] <- NA
month_3 = left_join(month_3,ref_month3, by = c("ParticipantIdentifier","StepStartDate_ET"))%>%
   mutate(days_complete_task = StepStartDate_ET - TaskInsertedDate)%>%
   filter(days_complete_task<=30) %>%   
   filter(IsDistanceAtLeast100m==1) 

month_6 = Valentine_Primary_Analysis_6MW_Data %>% filter(TaskIdentifier == "6 Minute Walk (6-mo.)")
ref_month6 = month_6 %>%
  group_by(ParticipantIdentifier)%>%
  summarise(TaskInsertedDate = min(TaskInsertedDate_ET),
            StepStartDate_ET = max(StepStartDate_ET,na.rm = T))
ref_month6[sapply(ref_month6, is.infinite)] <- NA
month_6 = left_join(month_6,ref_month6, by = c("ParticipantIdentifier","StepStartDate_ET"))%>%
   mutate(days_complete_task = StepStartDate_ET - TaskInsertedDate)%>%
   filter(days_complete_task<=30) %>%   
   filter(IsDistanceAtLeast100m==1) 


colnames(month_0)[19]="Distance_m_0"
colnames(month_3)[19]="Distance_m_3" 
colnames(month_6)[19]="Distance_m_6" 

```

```{r eval=FALSE, include=FALSE}

### Tables 

############0 month

nrow(month_0 %>% filter(days_complete_task<=7))

month_0 = Valentine_Primary_Analysis_6MW_Data %>% filter(TaskIdentifier == "6 Minute Walk (0-mo.)")
ref_month0 = month_0 %>%
  group_by(ParticipantIdentifier)%>%
  summarise(TaskInsertedDate = min(TaskInsertedDate_ET),
            StepStartDate_ET = max(StepStartDate_ET,na.rm = T))
ref_month0[sapply(ref_month0, is.infinite)] <- NA
month_0 = left_join(month_0,ref_month0, by = c("ParticipantIdentifier","StepStartDate_ET"))%>%
   mutate(days_complete_task = StepStartDate_ET - TaskInsertedDate)

df = month_0[!is.na(month_0$Distance_m),]
nrow(df)
nrow(df %>% filter(IsDistanceAtLeast100m!=1))
df = month_0[,c(4,34)]%>% group_by(ParticipantIdentifier) %>% summarise(days = max(days_complete_task,na.rm = T),
                                                                   reassigned = n() )
nrow(df %>% filter(reassigned>1))
nrow(df %>% filter(reassigned>1)%>% filter(days <= 30))

########### 3 month

nrow(month_3 %>% filter(days_complete_task<=7))

month_3 = Valentine_Primary_Analysis_6MW_Data %>% filter(TaskIdentifier == "6 Minute Walk (3-mo.)")
ref_month3 = month_3 %>%
  group_by(ParticipantIdentifier)%>%
  summarise(TaskInsertedDate = min(TaskInsertedDate_ET),
            StepStartDate_ET = max(StepStartDate_ET,na.rm = T))
ref_month3[sapply(ref_month3, is.infinite)] <- NA
month_3 = left_join(month_3,ref_month3, by = c("ParticipantIdentifier","StepStartDate_ET"))%>%
   mutate(days_complete_task = StepStartDate_ET - TaskInsertedDate)



df = month_3[!is.na(month_3$Distance_m),]
nrow(df)
nrow(df %>% filter(IsDistanceAtLeast100m!=1))
df = month_3[,c(4,34)]%>% group_by(ParticipantIdentifier) %>% summarise(days = max(days_complete_task,na.rm = T),
                                                                   reassigned = n() )
nrow(df %>% filter(reassigned>1))
nrow(df %>% filter(reassigned>1)%>% filter(days <= 30))



############# 6 month
nrow(month_6 %>% filter(days_complete_task<=7))

month_6 = Valentine_Primary_Analysis_6MW_Data %>% filter(TaskIdentifier == "6 Minute Walk (6-mo.)")
ref_month6 = month_6 %>%
  group_by(ParticipantIdentifier)%>%
  summarise(TaskInsertedDate = min(TaskInsertedDate_ET),
            StepStartDate_ET = max(StepStartDate_ET,na.rm = T))
ref_month6[sapply(ref_month6, is.infinite)] <- NA
month_6 = left_join(month_6,ref_month6, by = c("ParticipantIdentifier","StepStartDate_ET"))%>%
   mutate(days_complete_task = StepStartDate_ET - TaskInsertedDate)

df = month_6[!is.na(month_6$Distance_m),]
nrow(df)
nrow(df %>% filter(IsDistanceAtLeast100m!=1))
df = month_6[,c(4,34)]%>% group_by(ParticipantIdentifier) %>% summarise(days = max(days_complete_task,na.rm = T),
                                                                   reassigned = n() )
nrow(df %>% filter(reassigned>1))
nrow(df %>% filter(reassigned>1)%>% filter(days <= 30))
```


```{r include=FALSE}


Data_6MW = month_0[,c("ParticipantIdentifier","Distance_m_0")] %>% 
  full_join(month_3[,c("ParticipantIdentifier","Distance_m_3")], by = "ParticipantIdentifier" )%>% 
  full_join(month_6[,c("ParticipantIdentifier","Distance_m_6")],by = "ParticipantIdentifier" )%>%
  distinct(.keep_all = TRUE)

# ID_220 = unique(StudyParticipants$ParticipantIdentifier)  
# ID_169 = unique(Data_6MW[!is.na(Data_6MW$Distance_dif6),]$ParticipantIdentifier)
# ID = setdiff(ID_220,ID_169)




##### Consort diagram
# Valentine_Primary_Analysis_6MW_Data_wrong = Valentine_Primary_Analysis_6MW_Data %>%
#   filter(ParticipantIdentifier %in% ID)%>%
#   filter(TaskIdentifier != "6 Minute Walk (3-mo.)") %>%
#   left_join(StudyParticipants[,c("ParticipantIdentifier","Cohort")])
# 
# # i_d = c(97,872,568,935,847,434)
# # who don't have any data
# # ID = setdiff(ID,i_d)
# 
# wrong_summary = Valentine_Primary_Analysis_6MW_Data_wrong %>%
#   select(ParticipantIdentifier,Phase, TaskRank, Distance_m, Duration_min, Cohort)%>%
#   arrange(ParticipantIdentifier)%>%
#   mutate(Is_less100m = if_else(Distance_m>100,0,1),
#          IsDistance_NA = if_else(is.na(Distance_m),1,0))

##### Join with case/control

Data_6MW = left_join(Data_6MW, StudyParticipants[,c("ParticipantIdentifier","Cohort","Device","CardiacRehabCenter","IsExit","Gender", "Age_cat","Age_cat_id", "Race", "Indication",  "ComorbidCAD_bool", "ComorbidHF_bool", "ComorbidPCICABG_bool","ComorbidValve_bool")], by = "ParticipantIdentifier")

Data_6MW$Cohort = factor(Data_6MW$Cohort, levels = c("Usual","Telehealth"))
Data_6MW$Cohort = relevel(Data_6MW$Cohort, ref = "Usual")


# Summary

Data_6MW_summary = Data_6MW %>%
  group_by(Cohort) %>%
  summarise(D_0 = mean(Distance_m_0, na.rm = T),
            D_3 = mean(Distance_m_3, na.rm = T),
            D_6 = mean(Distance_m_6, na.rm = T),
            vD_0 = sd(Distance_m_0, na.rm = T),
            vD_3 = sd(Distance_m_3, na.rm = T),
            vD_6 = sd(Distance_m_6, na.rm = T))


Data_6MW_summary = data.frame(Cohort = rep(c("Usual","Telehealth"),3), 
                              Phase = rep(c("0-mo.","3-mo.","6-mo."), each = 2),
                              Distance = as.vector(as.matrix(Data_6MW_summary[,2:4])),
                              SE = as.vector(as.matrix(Data_6MW_summary[,5:7])))
```


```{r eval=FALSE, include=FALSE}
# ## Case
apply(Data_6MW[Data_6MW$Cohort == "Telehealth",c(2,3,4)], 2, mean, na.rm = T)
apply(Data_6MW[Data_6MW$Cohort == "Telehealth",c(2,3,4)], 2, sd, na.rm = T)
apply(!is.na(Data_6MW[Data_6MW$Cohort == "Telehealth",c(2,3,4)]), 2, sum)
## Control
apply(Data_6MW[Data_6MW$Cohort == "Usual",c(2,3,4)], 2, mean, na.rm = T)
apply(Data_6MW[Data_6MW$Cohort == "Usual",c(2,3,4)], 2, sd, na.rm = T)
apply(!is.na(Data_6MW[Data_6MW$Cohort == "Usual",c(2,3,4)]), 2, sum)
# 
# 
# ################ Supplement
# 
# 
## Case Apple
apply(Data_6MW[Data_6MW$Cohort == "Telehealth"& Data_6MW$Device == "Apple Watch",c(2,3,4)], 2, mean, na.rm = T)
apply(Data_6MW[Data_6MW$Cohort == "Telehealth"& Data_6MW$Device == "Apple Watch",c(2,3,4)], 2, sd, na.rm = T)
apply(!is.na(Data_6MW[Data_6MW$Cohort == "Telehealth"& Data_6MW$Device == "Apple Watch",c(2,3,4)]), 2, sum)


## Case Fitbit
apply(Data_6MW[Data_6MW$Cohort == "Telehealth"& Data_6MW$Device == "Fitbit",c(2,3,4)], 2, mean, na.rm = T)
apply(Data_6MW[Data_6MW$Cohort == "Telehealth"& Data_6MW$Device == "Fitbit",c(2,3,4)], 2, sd, na.rm = T)
apply(!is.na(Data_6MW[Data_6MW$Cohort == "Telehealth"& Data_6MW$Device == "Fitbit",c(2,3,4)]), 2, sum)
# 
# 
# 
## Control Apple

apply(Data_6MW[Data_6MW$Cohort == "Usual"& Data_6MW$Device == "Apple Watch",c(2,3,4)], 2, mean, na.rm = T)
apply(Data_6MW[Data_6MW$Cohort == "Usual"& Data_6MW$Device == "Apple Watch",c(2,3,4)], 2, sd, na.rm = T)
apply(!is.na(Data_6MW[Data_6MW$Cohort == "Usual"& Data_6MW$Device == "Apple Watch",c(2,3,4)]), 2, sum)

# 
# 
## Control Fibit

apply(Data_6MW[Data_6MW$Cohort == "Usual"& Data_6MW$Device == "Fitbit",c(2,3,4)], 2, mean, na.rm = T)
apply(Data_6MW[Data_6MW$Cohort == "Usual"& Data_6MW$Device == "Fitbit",c(2,3,4)], 2, sd, na.rm = T)
apply(!is.na(Data_6MW[Data_6MW$Cohort == "Usual"& Data_6MW$Device == "Fitbit",c(2,3,4)]), 2, sum)


```


### Plot 6MW distance 


```{r echo=FALSE, warning=FALSE}
####################
# plot and tables
####################

plot_multi_histogram(Valentine_Primary_Analysis_6MW_Data, "Distance_m", "TaskIdentifier")


### In the paragraph

table1(~Distance_m_0+Distance_m_3+Distance_m_6|Cohort, data = Data_6MW, overall=F, extra.col=list(`P-value`=pvalue) )
# 
# mean(Data_6MW$Distance_m_0)
# sd(Data_6MW$Distance_m_0)

# Valentine_Primary_Analysis_6MW_Data = Valentine_Primary_Analysis_6MW_Data %>%
#   left_join(StudyParticipants[,c("ParticipantIdentifier","Cohort")])

Valentine_Primary_Analysis_6MW_Data = Data_6MW %>% pivot_longer(cols =Distance_m_0: Distance_m_6,
                                                                names_to = "Phase", 
                                                                values_to = "Distance_m")

Valentine_Primary_Analysis_6MW_Data$Phase = if_else(Valentine_Primary_Analysis_6MW_Data$Phase=="Distance_m_0","0-mo.", if_else(Valentine_Primary_Analysis_6MW_Data$Phase=="Distance_m_3","3-mo.","6-mo."))

# Valentine_Primary_Analysis_6MW_Data$Cohort = if_else(Valentine_Primary_Analysis_6MW_Data$Cohort == "Usual", "Control", "Intervention")

# main plot

p_6mw_1 = ggplot(Valentine_Primary_Analysis_6MW_Data[Valentine_Primary_Analysis_6MW_Data$Phase!="3-mo.",])+
  geom_boxplot( aes(x = Phase, y = Distance_m, fill =Cohort))+
  facet_wrap(vars(Device))+
  ylab("6-minute Walk Distance (m)")+theme_bw()

p_6mw_1

p_6mw_2 = ggplot(Valentine_Primary_Analysis_6MW_Data[Valentine_Primary_Analysis_6MW_Data$Phase!="6-mo.",])+
  geom_boxplot( aes(x = Phase, y = Distance_m, fill =Cohort))+
  facet_wrap(vars(Device))+
  ylab("6-minute Walk Distance (m)")+theme_bw()

p_6mw_2

p_6mw_3 = ggplot(Valentine_Primary_Analysis_6MW_Data)+
  geom_boxplot( aes(x = Phase, y = Distance_m, fill =Cohort))+
  facet_wrap(vars(Device))+
  ylab("6-minute Walk Distance (m)")+theme_bw()

p_6mw_3

```


```{r echo=FALSE, warning=FALSE}
p = ggarrange(p_6mw_2,p_step_1,common.legend = TRUE,legend = "bottom")
```


```{r eval=FALSE, include=FALSE}
########### Comparison

with(Data_6MW, t.test(Distance_m_0[Cohort == "Usual"], Distance_m_0[Cohort == "Telehealth"])) # 0.5489
with(Data_6MW, t.test(Distance_m_3[Cohort == "Usual"], Distance_m_3[Cohort == "Telehealth"])) # 0.1778
with(Data_6MW, t.test(Distance_m_6[Cohort == "Usual"], Distance_m_6[Cohort == "Telehealth"])) # 0.3626


t.test(Data_6MW$`Distance_m_0`,Data_6MW$`Distance_m_3`) #0.1442
t.test(Data_6MW$`Distance_m_3`,Data_6MW$`Distance_m_6`) #0.7431
t.test(Data_6MW$`Distance_m_0`,Data_6MW$`Distance_m_6`) #0.3008
```



#### check the 6MW distance difference and distribution

```{r warning=FALSE}


Data_6MW$Distance_dif3 = Data_6MW$`Distance_m_3` - Data_6MW$`Distance_m_0`

hist(Data_6MW$Distance_dif3,breaks = 30)
summary(Data_6MW$Distance_dif3)

Data_6MW$Distance_dif6 = Data_6MW$`Distance_m_6` - Data_6MW$`Distance_m_0`

hist(Data_6MW$Distance_dif6,breaks = 30)
summary(Data_6MW$Distance_dif6)

```

```{r echo=FALSE, warning=FALSE}
########################
# table 2
#########################

cat("Baseline vs 6 month difference between Usual and Telehealth")

with(Data_6MW, t.test(Distance_dif6[Cohort == "Usual"], Distance_dif6[Cohort == "Telehealth"]))

plot_multi_histogram(Data_6MW, "Distance_dif6", "Cohort")


cat("Baseline vs 3 month difference between Usual and Telehealth")

with(Data_6MW, t.test(Distance_dif3[Cohort == "Usual"], Distance_dif3[Cohort == "Telehealth"])) 


plot_multi_histogram(Data_6MW, "Distance_dif3", "Cohort")

```


### The primary question to answer


```{r echo=FALSE}

fit_3 = lm(Distance_dif3 ~ Cohort * Device, data =Data_6MW )
summary(fit_3)

linearHypothesis(fit_3,c("CohortTelehealth = 0", "CohortTelehealth + CohortTelehealth:DeviceFitbit = 0"))

linearHypothesis(fit_3,"CohortTelehealth = 0")

linearHypothesis(fit_3,"CohortTelehealth + CohortTelehealth:DeviceFitbit = 0")


cat("---------------------------------","\n")

cat("There are intotal",dim(fit_3[["model"]])[1], "observations used for the baseline vs 3 months analysis","\n")

cat("---------------------------------","\n")

fit_6 = lm(Distance_dif6 ~ Cohort * Device, data =Data_6MW )
summary(fit_6)


cat("---------------------------------","\n")

cat("There are intotal",dim(fit_6[["model"]])[1], "observations used for the baseline vs 6 months analysis","\n")

cat("---------------------------------","\n")

linearHypothesis(fit_6,c("CohortTelehealth = 0", "CohortTelehealth + CohortTelehealth:DeviceFitbit = 0")) #0.1753
```


### Sub group analysis

```{r include=FALSE}

library("multcomp")
library(readxl)
library(tibble)
# library(forestplot)

dat_apple <-  read_excel("forestplotdata.xlsx")

dat_fitbit <- dat_apple


```

#### Age cat

For Apple Watch

```{r echo=FALSE}
fit_6 = lm(Distance_dif6 ~ Cohort *Age_cat, data = Data_6MW[Data_6MW$Device=="Apple Watch",] )
summary(fit_6)


dat_apple$`P Value`[1] = summary(fit_6)[["coefficients"]][4,4]
dat_apple[2,2:4]= t(tibble(c(coef(fit_6)[2],confint(fit_6)[2,])))
mod.lh = glht(fit_6, linfct = c("CohortTelehealth + CohortTelehealth:Age_catyoung = 0"))
dat_apple[3,2:4]= t(tibble(c(confint(mod.lh)[["confint"]])))
```


For Fitbit

```{r echo=FALSE}
fit_6 = lm(Distance_dif6 ~ Cohort *Age_cat, data = Data_6MW[Data_6MW$Device=="Fitbit",] )
summary(fit_6)

dat_fitbit$`P Value`[1] = summary(fit_6)[["coefficients"]][4,4]
dat_fitbit[2,2:4]= t(tibble(c(coef(fit_6)[2],confint(fit_6)[2,])))
mod.lh = glht(fit_6, linfct = c("CohortTelehealth + CohortTelehealth:Age_catyoung = 0"))
dat_fitbit[3,2:4]= t(tibble(c(confint(mod.lh)[["confint"]])))
```

#### Gender

For Apple Watch

```{r echo=FALSE}
fit_6 = lm(Distance_dif6 ~ Cohort *Gender, data = Data_6MW[Data_6MW$Device=="Apple Watch",] )
summary(fit_6)

dat_apple$`P Value`[5] = summary(fit_6)[["coefficients"]][4,4]
dat_apple[6,2:4]= t(tibble(c(coef(fit_6)[2],confint(fit_6)[2,])))
mod.lh = glht(fit_6, linfct = c("CohortTelehealth + CohortTelehealth:GenderM = 0"))
dat_apple[7,2:4]= t(tibble(c(confint(mod.lh)[["confint"]])))
```


For Fitbit

```{r echo=FALSE}
fit_6 = lm(Distance_dif6 ~ Cohort *Gender, data = Data_6MW[Data_6MW$Device=="Fitbit",] )
summary(fit_6)

dat_fitbit$`P Value`[5] = summary(fit_6)[["coefficients"]][4,4]
dat_fitbit[6,2:4]= t(tibble(c(coef(fit_6)[2],confint(fit_6)[2,])))
mod.lh = glht(fit_6, linfct = c("CohortTelehealth + CohortTelehealth:GenderM = 0"))
dat_fitbit[7,2:4]= t(tibble(c(confint(mod.lh)[["confint"]])))
```

#### Presence of Heart Failure

For Apple Watch

```{r echo=FALSE}
fit_6 = lm(Distance_dif6 ~ Cohort *ComorbidHF_bool , data = Data_6MW[Data_6MW$Device=="Apple Watch",] )
summary(fit_6)

dat_apple$`P Value`[13] = summary(fit_6)[["coefficients"]][4,4]
dat_apple[14,2:4]= t(tibble(c(coef(fit_6)[2],confint(fit_6)[2,])))
mod.lh = glht(fit_6, linfct = c("CohortTelehealth + CohortTelehealth:ComorbidHF_boolTRUE = 0"))
dat_apple[15,2:4]= t(tibble(c(confint(mod.lh)[["confint"]])))
```


For Fitbit

```{r echo=FALSE}
fit_6 = lm(Distance_dif6 ~ Cohort *ComorbidHF_bool, data = Data_6MW[Data_6MW$Device=="Fitbit",] )
summary(fit_6)

dat_fitbit$`P Value`[13] = summary(fit_6)[["coefficients"]][4,4]
dat_fitbit[14,2:4]= t(tibble(c(coef(fit_6)[2],confint(fit_6)[2,])))
mod.lh = glht(fit_6, linfct = c("CohortTelehealth + CohortTelehealth:ComorbidHF_boolTRUE = 0"))
dat_fitbit[15,2:4]= t(tibble(c(confint(mod.lh)[["confint"]])))
```

#### Cardiac Rehab Center

For Apple Watch

```{r echo=FALSE}
fit_6 = lm(Distance_dif6 ~ Cohort *CardiacRehabCenter, data = Data_6MW[Data_6MW$Device=="Apple Watch",] )
summary(fit_6)

dat_apple$`P Value`[9] = summary(fit_6)[["coefficients"]][4,4]
dat_apple[10,2:4]= t(tibble(c(coef(fit_6)[2],confint(fit_6)[2,])))

mod.lh = glht(fit_6, linfct = matrix(c(0,1,0,1),nrow = 1))
dat_apple[11,2:4]= t(tibble(c(confint(mod.lh)[["confint"]])))
```


For Fitbit

```{r echo=FALSE}
fit_6 = lm(Distance_dif6 ~ Cohort *CardiacRehabCenter, data = Data_6MW[Data_6MW$Device=="Fitbit",] )
summary(fit_6)

dat_fitbit$`P Value`[9] = summary(fit_6)[["coefficients"]][4,4]
dat_fitbit[10,2:4]= t(tibble(c(coef(fit_6)[2],confint(fit_6)[2,])))
mod.lh = glht(fit_6, linfct = matrix(c(0,1,0,1),nrow = 1))
dat_fitbit[11,2:4]= t(tibble(c(confint(mod.lh)[["confint"]])))
```


```{r eval=FALSE, include=FALSE}

## Apple Watch
subgps <- c(2,3,6,7,10,11,14,15)
dat_apple$Variable[subgps] <- paste("  ",dat_apple$Variable[subgps]) 

dat_apple = dat_apple %>% 
  mutate_if(is.numeric, round, digits = 2)

## Combine the count and percent column
np <- ifelse(!is.na(dat_apple$Low), paste("(",dat_apple$Low,",",dat_apple$High,")",sep=""), NA)
 
## The rest of the columns in the table. 
tabletext <- cbind(c("Subgroup","\n",dat_apple$Variable), 
                    c("95% CI","\n",np),
                    c("P Value","\n",dat_apple$`P Value`))



plot_apple <- forestplot(labeltext=tabletext, graph.pos=2, 
           mean=c(NA,NA,dat_apple$`Point Estimate`), 
           lower=c(NA,NA,dat_apple$Low), upper=c(NA,NA,dat_apple$High),
           title="Treatment Effect on 6MW Distance for Apple Watch Users",
           xlab="     <---Control Better---                                        ---Intervention Better--->",
           # xticks = NULL,
           txt_gp=fpTxtGp(label=gpar(cex=1.1),
                               ticks=gpar(cex=1.1),
                              xlab=gpar(cex = 1.2),
                              title=gpar(cex = 1.2)),
           col=fpColors(box="black", lines="black", zero = "gray50"),
           zero=0, cex=0.9, lineheight = "auto", boxsize=0.5, colgap=unit(6,"mm"),
           lwd.ci=2, ci.vertices=TRUE, ci.vertices.height = 0.4)

plot_apple


## Fitbit

dat_fitbit$Variable[subgps] <- paste("  ",dat_fitbit$Variable[subgps]) 

dat_fitbit = dat_fitbit %>% 
  mutate_if(is.numeric, round, digits = 2)

## Combine the count and percent column
np <- ifelse(!is.na(dat_fitbit$Low), paste("(",dat_fitbit$Low,",",dat_fitbit$High,")",sep=""), NA)
 
## The rest of the columns in the table. 
tabletext <- cbind(c("Subgroup","\n",dat_fitbit$Variable), 
                    c("95% CI","\n",np),
                    c("P Value","\n",dat_fitbit$`P Value`))



plot_fitbit <- forestplot(labeltext=tabletext, graph.pos=2, 
           mean=c(NA,NA,dat_fitbit$`Point Estimate`), 
           lower=c(NA,NA,dat_fitbit$Low), upper=c(NA,NA,dat_fitbit$High),
           title="Treatment Effect on 6MW Distance for Fitbit Users",
           xlab="     <---Control Better---                                        ---Intervention Better--->",
           txt_gp=fpTxtGp(label=gpar(cex=1.1),
                              ticks=gpar(cex=1.1),
                              xlab=gpar(cex = 1.2),
                              title=gpar(cex = 1.2)),
           col=fpColors(box="black", lines="black", zero = "gray50"),
           zero=0, cex=0.9, lineheight = "auto", boxsize=0.5, colgap=unit(6,"mm"),
           lwd.ci=2, ci.vertices=TRUE, ci.vertices.height = 0.4)

plot_fitbit

# plot <- ggplot(dat, aes(y = Index, x = OR)) +
#   geom_point(shape = 18, size = 5) +  
#   geom_errorbarh(aes(xmin = LL, xmax = UL), height = 0.25) +
#   geom_vline(xintercept = 1, color = "red", linetype = "dashed", cex = 1, alpha = 0.5) +
#   scale_y_continuous(name = "", breaks=1:8, labels = dat$label, trans = "reverse") +
#   xlab("6-minute walking distance") + 
#   ylab(" ") + 
#   theme_bw() +
#   theme(panel.border = element_blank(),
#         panel.background = element_blank(),
#         panel.grid.major = element_blank(), 
#         panel.grid.minor = element_blank(), 
#         axis.line = element_line(colour = "black"),
#         axis.text.y = element_text(size = 12, colour = "black"),
#         axis.text.x.bottom = element_text(size = 12, colour = "black"),
#         axis.title.x = element_text(size = 12, colour = "black"))+
#   facet_wrap(vars(Device))
# plot

```



### 3 month device, cohort effect

```{r}
fit_3_apple = lm(Distance_dif3 ~ Cohort, data = Data_6MW[Data_6MW$Device=="Apple Watch",])
summary(fit_3_apple)

confint(fit_3_apple)

fit_3_fitbit = lm(Distance_dif3 ~ Cohort, data = Data_6MW[Data_6MW$Device=="Fitbit",])
summary(fit_3_fitbit)

confint(fit_3_fitbit)
```




